---
jupyter:
  jupytext:
    formats: ipynb,Rmd
    text_representation:
      extension: .Rmd
      format_name: rmarkdown
      format_version: '1.2'
      jupytext_version: 1.9.1
  kernelspec:
    display_name: R
    language: R
    name: ir
---

# Setup

## Load libraries

```{r}
if (!require("xfun")) install.packages("xfun")

pkgs <- c(
  'urca', 'uroot', 'vars', 'forecast', 'dynlm', 'tsDyn', "aod", 
  'lubridate', 'tsbox', 'timetk', 'xts', 'tsibble', 'mltools',
  'tidyverse', 'readr', 'vroom', 'readxl', 'writexl', 
  'rlist', 'here', 'forcats', "janitor", 'data.table', 
  'extrafont', 'patchwork', 'hrbrthemes', 'ggthemes', 'ggsci', 'scales',
  'kableExtra', 'gt', 'reshape2'
)
xfun::pkg_attach2(pkgs, message = TRUE)

loadfonts(device = "win", quiet = TRUE)
setwd(here())

# source("R/granger_ty.R")
source("R/tsa.R")
```

## Load data

```{r}
price <- readRDS('data/rice/clean/rice_price.rds') %>%
  filter(date >= '2004-01-01') %>%
  mutate(
    across(phf:pwx, log),
    month = factor(month(date))
  )

price <- one_hot(as.data.table(price)) %>% 
  as_tsibble(index = date) %>% 
  select(-month_12)

price
```

```{r}
# price <- price %>%
#   mutate(
#     d_foodcrisis = ifelse(date >= '2007-11-01' & date <= '2008-06-01', 1, 0),
#     d_pledge = ifelse(date >= '2011-09-01' & date <= '2014-01-01', 1, 0)
#   )
```

# Hommali

```{r}
endo_h <- price %>% as_tibble() %>% select(phf, phw, phx)
exo_h  <- price %>% as_tibble() %>% select(d_foodcrisis, d_pledge, starts_with("month"))
```

## Descriptive Analysis

```{r}
endo_h %>% exp() %>%
  psych::describe() %>%
  rownames_to_column("varname") %>%
  select(varname, mean, sd, min, max) %>%
  write_xlsx("results/tbl_1_desc_h.xlsx")
```

```{r}
options(
  repr.plot.width = 10,
  repr.plot.height = 6 
)

price %>% 
  ggplot(aes(x=date)) +
    geom_line( aes(y=phf %>% exp(), color = 'เกษตรกรขายได้')) +
    geom_line( aes(y=phw %>% exp(), color = 'ขายส่ง')) +
    geom_line( aes(y=phx %>% exp(), color = 'ส่งออก')) +
    scale_color_d3() +
    expand_limits(y = 0) +
    scale_y_continuous(label = comma) +
    labs(title = 'ราคาข้าวหอมมะลิ 3 ตลาด',
         x =  "ปี",
         y = 'บาท/ตัน'
        ) +
    theme_ipsum(
      base_size = 16,
      base_family = "Athiti Light",
      axis_title_size = 16
    ) +
    theme(
      legend.position = c(1, 1),
      legend.direction = 'horizontal',
      legend.justification = 'right',
      legend.title = element_blank()
    ) +
    scale_x_date(date_breaks = '2 years',
                 labels = function(x) year(x)+543
                 )

# ggsave('figures/fig_1_h.png', width = 10, height = 6)
```

## Unit Root Test

```{r}
# urca::ur.df(
#   price %>% select('pwx') %>% ts_ts %>% diff(), 
#   type = 'drift') %>% summary
```

```{r}
ph_names <- c('phf', 'phw', 'phx')

expand_grid(
  'varname' = ph_names, 
  'n_diff' = c(0,1)
) %>% 
  pmap(tsa_ur_adf, y = price, specs = c('trend', 'drift')) %>%
  reduce(bind_rows) %>% 
  tsa_report_adf() %>% 
  write_xlsx("results/tbl_2_unit_root_h.xlsx")
```

## Optimal Lag

```{r}
varselect <- VARselect(endo_h, exogen = exo_h, lag.max = 10, type = 'both')
varselect
```

```{r}
varselect$criteria %>% t %>% 
  as_tibble %>% 
  rownames_to_column('lag') %>%
  write_xlsx("results/tbl_3_optimal_lag_h.xlsx")
```

## Deterministic Term

```{r}
tsa_select_DT <- function(endo, ...) {
  
  # estimate 5 specification of VECM
  
  spec1 <- tsDyn::VECM(endo, estim = "ML", include = 'none', ...) %>% suppressWarnings()
  spec2 <- tsDyn::VECM(endo, estim = "ML", LRinclude = 'const', ...) %>% suppressWarnings()
  spec3 <- tsDyn::VECM(endo, estim = "ML", include = 'const', ...) %>% suppressWarnings()
  spec4 <- tsDyn::VECM(endo, estim = "ML", include = 'const', LRinclude = 'trend', ...) %>% suppressWarnings()
  spec5 <- tsDyn::VECM(endo, estim = "ML", include = 'both', ...) %>% suppressWarnings()
  
  res <- list(spec1, spec2, spec3, spec4, spec5)
  tibble(
    model = 1:length(res),
    AIC = res %>% map_dbl(AIC),
    BIC = res %>% map_dbl(BIC)
  )
}

tsa_select_DT(endo_h, exogen = exo_h, lag = 1) %>%
  write_xlsx("results/tbl_a1_model_spec.xlsx")
```

## Cointegration Test

```{r}
vecm_h <- tsDyn::VECM(
  endo_h, exogen = exo_h, lag = 1,
  estim = "ML", LRinclude = 'const',
  ) %>% suppressWarnings()
```

```{r}
summary(vecm_h)$coefMat %>% as.data.frame %>% 
  rownames_to_column() %>%
  write_xlsx("results/vecm_ph.xlsx")
```

```{r}
vecm_h <- tsDyn::VECM(
  endo_h, exogen = exo_h, lag = 1,
  estim = "ML", LRinclude = 'const',
  ) %>% suppressWarnings()

rank.test(vecm_h) %>% summary() %>%
  select(-trace_pval_T) %>%
  write_xlsx("results/tbl_4_rank_test.xlsx")
```

## VECM

```{r}
vecm_h <- tsDyn::VECM(
  endo_h, exogen = exo_h, lag = 1, r = 2,
  estim = "ML", LRinclude = 'const',
  ) %>% suppressWarnings()
```

```{r}
vecm_h %>% summary
```

```{r}
vecm_h$model.specific$beta
```

```{r}
# vecm_h %>% summary %>% suppressWarnings()
```

```{r}
vecm_h %>% 
  summary %>% suppressWarnings() %>% 
  '[['('bigcoefficients') %>% 
  t %>% 
  as.data.frame %>%
  rownames_to_column() %>%
  write_xlsx("results/tbl_5_vecm_h.xlsx")
```

```{r}
# vecm_h_cajo <- ca.jo(endo_h, dumvar = exo_h, ecdet = "const", type="eigen", K=2, spec="transitory")
# vecm_h_ols <- cajools(vecm_h_cajo)
# vecm_h_ols %>% summary()
```

### Test

```{r}
vecm_h_var <- tsDyn:::vec2var.tsDyn(vecm_h)
tsa_VAR_diag(vecm_h_var) %>%
  write_xlsx("results/tbl_a2_model_diag_h.xlsx")
```

```{r}
bound_unit_circle <- function(x) {
  x <- ifelse(x > 1, 1, x)
  x <- ifelse(x < -1, -1, x)
  return(x)
}

# bound_unit_circle(Re(eigen(vecm_h_var$A$A1)$values))
```

```{r}
library("plotrix")

options(
  repr.plot.width = 6,
  repr.plot.height = 6 
)

png(
  filename = "figures/fig_a1_stability_h.png",
  units = "in", 
  width = 6, 
  height = 6,
  pointsize = 10, 
  res = 300
)

x <- seq(-1, 1, length = 1000)
y1 <- sqrt(1-x^2)
y2 <- -sqrt(1-x^2)
plot(
  Re(eigen(vecm_h_var$A$A1)$values) %>% bound_unit_circle,
  Im(eigen(vecm_h_var$A$A1)$values) %>% bound_unit_circle,
  pch=19, col = "blue", 
  xlab = 'Real part', ylab = 'Complex part',
  ylim = c(-2, 2), xlim = c(-2, 2)
)
points(
  Re(eigen(vecm_h_var$A$A2)$values) %>% bound_unit_circle,
  Im(eigen(vecm_h_var$A$A2)$values) %>% bound_unit_circle,
  pch=19, col = "blue"
)
abline(h=0)
abline(v=0)
draw.circle(0, 0, 1, nv = 1000, border = NULL, col = NA, lty = 1, lwd = 1)

dev.off()
```

## Granger Casuality Test

```{r}
tsa_gc_ty(
  endo_h, exogen = exo_h, 
  p=1, m=1, type = "const"
) %>% select(-wald)
```

```{r}
tsa_gc_ty(
  endo_h, exogen = exo_h, 
  p=1, m=1, type = "const"
) %>% 
  select(-wald) %>%
  mutate(chi2_pval = round(chi2_pval, 5)) %>%
  write_xlsx("results/tbl_a3_granger_h.xlsx")
```

## IRF

```{r}
vecm_h <- tsDyn::VECM(
  endo_h %>% select(1,2,3), exogen = exo_h, lag = 1, r = 2,
  estim = "ML", LRinclude = 'const',
  ) %>% suppressWarnings()
```

```{r}
# tsDyn::irf(vecm_h, boot = TRUE, n.ahead = 36, ortho = TRUE)
```

```{r}
for (i in c("phf", "phw", "phx")) {
  png(
    filename = glue::glue("figures/fig_3_irf1_{i}.png"),
    units = "in", 
    width = 7, 
    height = 7,
    pointsize = 12, 
    res = 300
  )
  tsDyn::irf(vecm_h, impulse = i, boot = TRUE, n.ahead = 24, ortho = TRUE) %>% plot
  dev.off()  
}
```

# White Rice

```{r}
endo_w <- price %>% as_tibble() %>% select(pwf, pww, pwx)
exo_w  <- price %>% as_tibble() %>% select(d_foodcrisis, d_pledge)
```

```{r}

```

## Descriptive Analysis

```{r}
endo_w %>% exp() %>%
  psych::describe() %>% 
  rownames_to_column("varname") %>%
  select(varname, mean, sd, min, max) %>%
  write_xlsx("results/tbl_1_desc_w.xlsx")
```

```{r}
options(
  repr.plot.width = 10,
  repr.plot.height = 6 
)

price %>% 
  ggplot(aes(x=date)) +
    geom_line( aes(y=pwf %>% exp(), color = 'เกษตรกรขายได้')) +
    geom_line( aes(y=pww %>% exp(), color = 'ขายส่ง')) +
    geom_line( aes(y=pwx %>% exp(), color = 'ส่งออก')) +
    scale_color_d3() +
    expand_limits(y = 0) +
    scale_y_continuous(label = comma) +
    labs(title = 'ราคาข้าวเจ้าขาว 3 ตลาด',
         x =  "ปี",
         y = 'บาท/ตัน'
        ) +
    theme_ipsum(
      base_size = 16,
      base_family = "Athiti Light",
      axis_title_size = 16
    ) +
    theme(
      legend.position = c(1, 1),
      legend.direction = 'horizontal',
      legend.justification = 'right',
      legend.title = element_blank()
    ) +
    scale_x_date(date_breaks = '2 years',
                 labels = function(x) year(x)+543
                 )

ggsave('figures/fig_1_w.png', width = 10, height = 6)
```

## Unit Root Test

```{r}
p_w_names <- c('pwf', 'pww', 'pwx')

expand_grid(
  'varname' = p_w_names, 
  'n_diff' = c(0,1)
) %>% 
  pmap(tsa_ur_adf, y = price) %>%
  reduce(bind_rows) %>% 
  tsa_report_adf() %>% 
  write_xlsx("results/tbl_2_unit_root_w.xlsx")
```

## Optimal Lag

```{r}
varselect <- VARselect(endo_w, exogen = exo_w, lag.max = 10, type = 'both')
varselect
```

```{r}
varselect$criteria %>% t %>% 
  as_tibble %>% 
  rownames_to_column('lag') %>%
  write_xlsx("results/tbl_3_optimal_lag_w.xlsx")
```

```{r}
VAR(endo_w, exogen = exo_w, p = 2, type = "both") %>% tsa_VAR_diag
```

## Test

```{r}
VAR(endo_w, exogen = exo_w, p = 2, type = "both") %>% 
  tsa_VAR_diag %>%
  write_xlsx("results/tbl_a2_model_diag_w2.xlsx")
```

```{r}
VAR(endo_w, exogen = exo_w, p = 6, type = "both") %>% 
  tsa_VAR_diag %>%
  write_xlsx("results/tbl_a2_model_diag_w6.xlsx")
```

```{r}
var_w <- tsDyn::lineVar(
  endo_w, exogen = exo_w, lag = 6, 
  model = "VAR", include = "both", r = 1
) %>% suppressWarnings()
```

```{r}
var_w %>% 
  summary %>% suppressWarnings() %>% 
  '[['('bigcoefficients') %>% 
  t %>% 
  as.data.frame %>%
  rownames_to_column() %>%
  write_xlsx("results/tbl_5_var_w.xlsx")
```

```{r}
# var_w %>% str()
```

```{r}
library("plotrix")

options(
  repr.plot.width = 6,
  repr.plot.height = 6 
)

png(
  filename = "figures/fig_a1_stability_w.png",
  units = "in", 
  width = 6, 
  height = 6,
  pointsize = 10, 
  res = 300
)

x <- seq(-1, 1, length = 1000)
y1 <- sqrt(1-x^2)
y2 <- -sqrt(1-x^2)
plot(
  Re(eigen(var_h_var$A$A1)$values) %>% bound_unit_circle,
  Im(eigen(vecm_h_var$A$A1)$values) %>% bound_unit_circle,
  pch=19, col = "blue", 
  xlab = 'Real part', ylab = 'Complex part',
  ylim = c(-2, 2), xlim = c(-2, 2)
)
points(
  Re(eigen(vecm_h_var$A$A2)$values) %>% bound_unit_circle,
  Im(eigen(vecm_h_var$A$A2)$values) %>% bound_unit_circle,
  pch=19, col = "blue"
)
abline(h=0)
abline(v=0)
draw.circle(0, 0, 1, nv = 1000, border = NULL, col = NA, lty = 1, lwd = 1)

dev.off()
```

```{r}
var_w <- VAR(endo_w, exogen = exo_w, p = 6, type = "both")
```

```{r}
var_w %>% summary
```

```{r}
# tsDyn::lineVar(endo_w, exogen = exo_w, lag = 6, model = "VAR", include = "both") %>% summary
```

## Granger Casuality Test

```{r}
tsa_gc_ty(
  endo_w, exogen = exo_w, 
  p = 6, m = 0, type = "both"
) %>% select(-wald)
```

```{r}
tsa_gc_ty(
  endo_w, exogen = exo_w, 
  p = 6, m = 0, type = "both"
) %>% 
  select(-wald) %>%
  mutate(chi2_pval = round(chi2_pval, 5)) %>%
  write_xlsx("results/tbl_a3_granger_w.xlsx")
```

## IRF

```{r}
for (i in c("pwf", "pww", "pwx")) {
  png(
    filename = glue::glue("figures/fig_3_irf_{i}.png"),
    units = "in", 
    width = 7, 
    height = 7,
    pointsize = 12, 
    res = 300
  )
  vars::irf(var_w, impulse = i, boot = TRUE, n.ahead = 24, ortho = TRUE) %>% plot
  dev.off()  
}
```

```{r}

```

```{r}

```

```{r}
vecm_wht <- tsDyn::linear(
  endo, estim = "ML", include = 'none', LRinclude = 'const', 
  lag = 2, exogen = exo, r = 1
) %>% suppressWarnings()
```

```{r}
lags = 8
pt.asympt = rep(0, lags)
pt.adj = rep(0, lags)
bg = rep(0, lags)
es = rep(0, lags)

for (i in 1:lags) {
  pt.asympt[i] = serial.test(wht_var, lags.pt = i, type = "PT.asymptotic")$serial$p.value
  pt.adj[i] = serial.test(wht_var, lags.pt = i, type = "PT.adjusted")$serial$p.value
  bg[i] = serial.test(wht_var, lags.bg = i, type = "BG")$serial$p.value
  es[i] = serial.test(wht_var, lags.bg = i, type = "ES")$serial$p.value
}

data.frame(pt.asymptotic = pt.asympt, pt.adjusted = pt.adj, BG = bg, ES = es)
```

```{r}
VAR(endo_w, exogen = exo_w, p = 6, type = "const") %>% summary
```

```{r}
granger_ty(
  endo , p = 2, m = 1, type = "const", exogen = exo
) %>% select(-wald)
```

```{r}
VARselect(endo_w, exogen = exo_w, lag.max = 10, type = 'both')

var_w <- VAR(endo_w, exogen = exo_w, p = 6, type = "both")
var_w %>% tsa_VAR_diag

var_w %>% summary
```

```{r}

```

```{r}
x <- rh %>% exp() %>%
  ts_data.frame %>%
  ts_wide %>%
  clean_names %>%
  mutate(
    ws2fg = rhws/rhfg,
    xb2fg = rhxb/rhfg,
    xb2ws = rhxb/rhws
  )
```

```{r}
describe(x)
```

```{r}
ggplot(x, aes(x = time, y = ws2fg)) +
  geom_line(color = "red") +
  geom_line(aes(x = time, y = xb2fg), color = "blue")
```

```{r}
x %>%
  filter(rhws > rhxb)
```

```{r}
ggplot(x, aes(x = time, y = rhws)) +
  geom_line(color = "red") +
  geom_line(aes(x = time, y = rhxb), color = "blue")
```

# Unit Root Test

```{r}
# get object name as a string
get_objname <- function(x) deparse(substitute(x))

# unit root test for a time series
tsf_ur_adf <- function(y, n_diff = 0, varname = NULL, ...) {
  
  varname <- if (is.null(varname)) deparse(substitute(y)) else varname
  y <- y %>% ts_ts
  y <- if(n_diff == 0) y else diff(y, n_diff)
  
  ur.trend <- ur.df(y, type='trend', lags = 10, selectlags = "BIC", ...)
  ur.drift <- ur.df(y, type='drift', lags = 10, selectlags = "BIC", ...)
  ur.none  <- ur.df(y, type='none' , lags = 10, selectlags = "BIC", ...)

  tstat.trend <- ur.trend@teststat
  tstat.drift <- ur.drift@teststat
  tstat.none  <- ur.none@teststat

  cv.trend <- ur.trend@cval
  cv.drift <- ur.drift@cval
  cv.none  <- ur.none@cval
  
  lags.trend <- dim(ur.trend@testreg$coefficients)[1] - 3
  lags.drift <- dim(ur.drift@testreg$coefficients)[1] - 2
  lags.none  <- dim(ur.none@testreg$coefficients)[1]  - 1

  df_test <- rbind(
    cbind(t(tstat.trend), cv.trend, lags.trend),
    cbind(t(tstat.drift), cv.drift, lags.drift),
    cbind(t(tstat.none) , cv.none , lags.none)
  ) %>% 
    as.data.frame() %>%
    rename(c("lag" = "lags.trend")) %>%
    rownames_to_column("hypo") %>%
    filter(str_starts(hypo, 'tau')) %>%
    mutate(
      result = ifelse(abs(statistic) >= abs(`5pct`), 'Reject', 'Accept'),
      variable = varname,
      level = paste0('d', n_diff),
      star = case_when(
        abs(statistic) > abs(`1pct`) ~ "***",
        abs(statistic) > abs(`5pct`) ~ "** ",
        abs(statistic) > abs(`10pct`) ~ "*  ",
        TRUE ~ "   "
      ),
      hypo = case_when(
        hypo == "tau3" ~ "trend",
        hypo == "tau2" ~ "constant",
        hypo == "tau1" ~ "none"
      ),
      statistic = format(round(statistic, 3), nsmall = 3) %>% 
        as.character %>% str_pad(6, "left")
    ) %>%
    select(variable, hypo, level, everything())
  
  return(df_test)
}
```

```{r}
# unit root test in a table format
tsf_report_adf <- function(adf_test) {
  adf_test %>% 
    mutate(
      stat_star = glue::glue(
        "{statistic}{star}({lag})"
      ) %>% as.character()
    ) %>%
    pivot_wider(
      id_cols = c("variable"),
      names_from = c("level", "hypo"),
      values_from = c("stat_star")
    )
}
```

```{r}
rh_names <- c('RHFG', 'RHWS', 'RHXB')
rw_names <- c('RWFG', 'RWWS', 'RWXB')

params_rh <- expand_grid(rh_names, c(0,1)) %>%
  set_names(c('name', 'ndiff')) %>%
  mutate(ser = rh[, name] %>% as.list)

params_rw <- expand_grid(rw_names, c(0,1)) %>%
  set_names(c('name', 'ndiff')) %>%
  mutate(ser = rw[, name] %>% as.list)

params <- rbind(params_rh, params_rw)
```

```{r}
library(psych)
s1 <- describe(x$rh$endo %>% exp %>% ts_wide) %>% as_tibble()
s2 <- describe(x$rw$endo %>% exp %>% ts_wide) %>% as_tibble()
s0 <- bind_rows(s1, s2) %>% select(mean, sd, min, max)
s0 %>% round(0)
```

# Steps


## Overall Procedure
1. Model Specification
  - Unit Root Test -> `urca::ur.df`
  - AR Order or Lag Selection -> `VARselect`
  - Functional Form -> `tsDyn::VECM` and `AIC`, `BIC`
  - Cointegrating Rank -> `tsDyn::rank.test`
  - Choice of Deterministic Term
    - exos + seasonal 

2. Model Checking
  - Descriptive Analysis
  - Diagnostic Tests
    * Serial Autocorrelation
    * Heteroskedasticity
    * Normality
    * Stability
    
3. Analysis
  - Interpret
  - Test economic theory
  - Granger Causality
  - IRF


## Toda Yamamoto Procedure
1. Stationary test and find maximum order of integration called `m`
2. Determine optimal lag of VAR in levels using information criteria called `p`
3. Given lag `p` estimate VAR(p) and test for serial correlation. If failed to pass add more lag until there is no serial correlation problem.
4. Conduct a cointegration test i.e. Johansen test keep result for cross-checking
5. Estimate VAR(p+m)
6. Wald test of the first `p` lags not `p+m` and associated test statistics is distributed as Chi-square(p)


# Vertical PT: Hommali Rice

```{r}
endo <- x$rh$endo
exo <- x$rh$exo
```

```{r}
plot(endo)
```

## Model Specification


### Lag Selection

```{r}
VARselect(endo, lag.max = 5, type = 'const')$criteria %>% t
```

```{r}
VARselect(endo, lag.max = 5, type = 'both')
```

```{r}
VARselect(endo, lag.max = 5, type = 'both')$criteria %>% t %>% 
  as_tibble() %>% janitor::clean_names() %>%
  select(4,1,3,2) %>%
  write_xlsx('results/02_lag_select_rh.xlsx')
```

```{r}
check_VAR <- function(var_obj) {
  res <- list()
  res$serial <- serial.test(var_obj, lags.pt = 12, type = 'PT.asymptotic')
  res$arch   <- arch.test(var_obj, lags.multi = 12)
  res$norm   <- normality.test(var_obj)
  return(res)
}

VAR(endo, p = 1, type = "both") %>% check_VAR()
```

> **optimal lag** = 1


### Deterministic Terms

```{r}
x <- VECM(Canada, lag = 2, estim = "ML", include = 'both',  LRinclude = 'both')
rank.test(x)
```

```{r}
select_DT <- function(endo, exo) {
  spec1 <- VECM(endo, lag = 1, estim = "ML", include = 'none',  LRinclude = 'none' , exogen = exo)
  spec2 <- VECM(endo, lag = 1, estim = "ML", include = 'none',  LRinclude = 'const', exogen = exo)
  spec3 <- VECM(endo, lag = 1, estim = "ML", include = 'const', LRinclude = 'const', exogen = exo)
  spec4 <- VECM(endo, lag = 1, estim = "ML", include = 'const', LRinclude = 'trend', exogen = exo)
  spec5 <- VECM(endo, lag = 1, estim = "ML", include = 'trend', LRinclude = 'both' , exogen = exo)
#   spec6 <- VECM(endo, lag = 1, estim = "ML", include = 'none',  LRinclude = 'trend' , exogen = exo)
  
  res <- list(spec1, spec2, spec3, spec4, spec5)
  tibble(
    model = 1:length(res),
    AIC = res %>% map_dbl(AIC),
    BIC = res %>% map_dbl(BIC)
  ) %>% arrange(BIC)
}

select_DT(endo, exo = exo)
```

```{r}
x <- VECM(endo, lag = 1, estim = "ML", include = 'none',  LRinclude = 'const', exogen = exo)
rank.test(x) %>% summary() #%>% 
#   write_xlsx('results/rh_03_rank_test.xlsx')
```

```{r}
x %>% summary()
```

```{r}
vecm_selected <- VECM(endo, lag = 1, estim = "ML", include = 'const', LRinclude = 'trend' , exogen = exo)
rank.test(vecm_selected)
```

```{r}
vecm_est <- VECM(endo, lag = 1, r = 2, estim = "ML", include = 'const', LRinclude = 'both', exogen = exo)
```

```{r}
vecm_est %>% summary()
```

```{r}
vecm_est$model.specific$beta
```

```{r}
var.rh <- VAR(endo, lag.max = 4, type = 'const', exogen = exo)
```

```{r}
vecm_rh.trace <- ca.jo(
  x$rh$endo, type = 'trace', ecdet = 'const', K = 2,
  spec = 'transitory', dumvar = x$rh$exo, season = 12
)

vecm_rh.eigen <- ca.jo(
  x$rh$endo, type = 'eigen', ecdet = 'const', K = 2, 
  spec = 'transitory', dumvar = x$rh$exo, season = 12
)

summary(vecm_rh.trace)
summary(vecm_rh.eigen)
```

> r = 2

```{r}
beta <- vecm_rh.trace@V
beta[, 2] <- beta[, 2] / beta[2,2]
beta[, 3] <- beta[, 3] / beta[3,3]
beta
```

```{r}
# tsDyn::VECM(endo, r = 2, lag = 1, exogen = exo, estim = "ML") %>% str
```

```{r}
vecm_rh <- cajorls(vecm_rh.trace, r = 2)
vecm_rh$rlm %>% summary
```

```{r}
vecm_rh
```

```{r}
# vecm_rh$rlm %>% summary
```

```{r}
21664.70*0.76
```

```{r}
21664.70*0.9107844
```

```{r jupyter={'outputs_hidden': True}}
endo
```

```{r}
1.55/0.76
```

```{r}
vecm_rh$beta %>% as.data.frame
```

```{r}
data(Canada)
vecm.x <- ca.jo(Canada)
var.x <- vec2var(vecm.x)
```

```{r}
var.x$A
```

## Model Checking

```{r}
Mod(eigen(diag(3) - var_rh$A$A1 - var_rh$A$A2)$values)
```

```{r}
var_rh <- vec2var(vecm_rh.trace, r = 2)
serial1 <- serial.test(var_rh, lags.pt = 12, type = 'PT.asymptotic')
arch1 <- arch.test(var_rh, lags.multi = 12)
norm1 <- normality.test(var_rh)

serial1$serial
arch1$arch.mul
norm1$jb.mul$JB
```

```{r}
vec2var
```

```{r}
eigen(var_rh$A$A1)$values
```

```{r}
eigen(var_rh$A$A2)
```

## Analysis


### Granger


### IRF


### LOP

```{r}

```

```{r}
H1 = ca.jo(
  x$rh$endo, type = 'trace', ecdet = 'const', K = 2, 
  spec = 'transitory', dumvar = x$rh$exo
)
H1 %>% summary()
```

```{r}
HA = matrix(c( 1, 0, 0,
               0, 1, 0,
              -1, 0, 0,
               0, 0, 1
             ), 
            ncol = 3, byrow = TRUE)

blrtest(z = vecm_rh.trace, H = HA, r = 1) %>% summary()
```

```{r}
save_irf_plot = function(model, impulse, width = 6, height = 8, n.ahead = 12) {
  model_name = deparse(substitute(model))
  filename = paste0("../img/irf_", model_name, '_', impulse, ".png")
  png(filename = filename,
    units = "in", 
    width = width, 
    height = height,
    pointsize = 10, 
    res = 300)
  plot(irf(model, impulse = impulse, n.ahead = n.ahead))
  dev.off()
}
```

```{r}
save_irf_plot(var_rh, 'RHFG')
save_irf_plot(var_rh, 'RHWS')
save_irf_plot(var_rh, 'RHXB')
```

```{r}
TVECM.HStest(endo[, c(1,3)])
```

```{r}
# TVECM(endo[, c(1,3)], common = 'only_ECT', include = 'const') %>% summary()
```

```{r}
# tsDyn::VECM(endo, lag = 1, r = 2, estim = "ML",
#             include = 'const', exogen = exo) %>% summary()
```

# Vertical : White Rice

```{r}
endo <- x$rw$endo
exo <- x$rw$exo
```

```{r}
summary(endo %>% exp())
```

```{r}

```

```{r}
# set_figsize(10, 6)
filename = paste0("../img/plot_", 'rhfg', ".png")
png(filename = filename,
  units = "in", 
  width = 12, 
  height = 6, 
  pointsize = 10, 
  res = 300)
plot.ts(endo)
dev.off()
```

## Model Specification

```{r}
VARselect(endo, lag.max = 5, type = 'const')
```

```{r}
VARselect(endo, lag.max = 5, type = 'both')
```

```{r}
VARselect(endo, lag.max = 5, type = 'const')$criteria %>% t %>% 
  as_tibble() %>% janitor::clean_names() %>%
  select(4,1,3,2) %>%
  write_xlsx('results/02_lag_select_rw.xlsx')
```

```{r}
vecm.trace = ca.jo(
  endo, type = 'trace', ecdet = 'const', K = 2,
  spec = 'transitory', dumvar = exo
)

vecm.eigen = ca.jo(
  endo, type = 'eigen', ecdet = 'const', K = 2, 
  spec = 'transitory', dumvar = exo
)

summary(vecm.trace)
summary(vecm.eigen)
```

### Law of One Price

```{r}
HA = matrix(c( 1, 1, 1, 1,
              -1, 0, 0, 0,
               0,-1, 0, 0,
               0, 0,-1, 0,
               0, 0, 0,-1
             ), 
            ncol = 4, byrow = TRUE)

blrtest(z = H1, H = HA, r = 4) %>% summary()

# H5 = matrix(c( 1, -1), 
#             ncol = 1, byrow = TRUE)

# H5 = c(1, -1)
# Hlop = matrix(c(1, -1, -1), nrow = 3, ncol = 1)

# bh5lrtest(z = H1, H = H5, r = 1) %>% summary()
```

```{r}
rh.cajolst = cajolst(x$rh$endo, trend = TRUE)
slotNames(rh.cajolst)
summary(rh.cajolst)
```

```{r}

```

```{r}
index(x$rh$endo)[59]
```

```{r}
rh.cajolst@bp
```

```{r}
rh.cajolst
```

```{r}
args(ca.jo)
```

```{r}
ca.jo(rh05, type = 'eigen', ecdet = "trend") %>% summary()
```

```{r}
summary(rh.cajolst)
```

```{r}
index(rw)[13]
```

# VAR


## Model Specification

```{r}
args(VARselect)
```

```{r}
VARselect(rh05, lag.max = 12, type = "both")
```

## Modle Estimation

```{r}
args(VAR)
```

```{r}
rh05.VAR = VAR(rh05, p = 2, type = "both")
```

```{r}
rh10.VAR = VAR(rh10, p = 2, type = "both")
```

```{r}
summary(rh05.VAR)
```

## Model Daignostic

```{r}
roots(rh05.VAR)
```

### Autocorrelation

```{r}
args(serial.test)
```

```{r}
rh05.VAR.ser = serial.test(rh05.VAR)
rh05.VAR.ser
```

```{r}

```

```{r}
set_figsize(10, 8)
rh05.VAR.ser = serial.test(rh05.VAR)
plot(rh05.VAR.ser, names = "RHFG")
plot(rh05.VAR.ser, names = "RHWS")
plot(rh05.VAR.ser, names = "RHXB")
```

### Hetero

```{r}
sp_vector <- c(1973.1,1961.53,1993.23,1937.27,2044.57,2054.27,2028.18,2082.2,2079.99,2094.86,2111.94,
               2099.29,2094.14,2039.87,1944.41,2024.81,2080.62,2054.08,1918.6,1904.42,2021.95,
               2075.54,2065.55,2083.89,2148.9,2170.95,2157.69,2143.02,2164.99,2246.63,2275.12,
               2329.91,2366.82,2359.31,2395.35,2433.99)
sp_ts <- ts(sp_vector,start=c(2014,7),frequency=12)
```

```{r}
dum = ts(rep(1, 13), start=c(2008, 10),frequency = 12)
```

```{r}
c(as.xts(rh05), as.xts(dum))
```

```{r}
data = replace_na(ts.union(rh05, dum), 0)
```

```{r}
# data
```

```{r}
VAR(data, p = 3, type = "both", exogen = dum)
```

```{r}

```

```{r}
rh05.VAR = VAR(rh05, p = 4, type = "both", season = 12)
```

```{r}
rh05.VAR.arch = arch.test(rh05.VAR, lags.multi = 12)
rh05.VAR.arch
```

```{r}
rh05.VAR %>% normality.test %>% plot
```

## Normality

```{r}
rh05.VAR %>% stability %>% plot
```

# Cointegration Analysis

```{r}
args(ca.jo)
```

```{r}
vars::VARselect(rh05)
vars::VARselect(rh06)
```

```{r}
rhrg  = price[, c('RHFG', 'RGFG')] %>% log() %>% na.omit()
```

```{r}
rh.eigen = ca.jo(rh, type = "eigen", K = 2, ecdet = "trend")
rh.trace = ca.jo(rh, type = "trace", K = 2, ecdet = "trend")
rw.eigen = ca.jo(rw, type = "eigen", K = 2, ecdet = "trend")
rw.trace = ca.jo(rw, type = "trace", K = 2, ecdet = "trend")
rg.eigen = ca.jo(rg, type = "eigen", K = 2, ecdet = "trend")
rg.trace = ca.jo(rg, type = "trace", K = 2, ecdet = "trend")

rhfg.eigen = ca.jo(rhfg, type = "eigen", K = 2, ecdet = "trend")
rhfg.trace = ca.jo(rhfg, type = "trace", K = 2, ecdet = "trend")
rwfg.eigen = ca.jo(rwfg, type = "eigen", K = 2, ecdet = "trend")
rwfg.trace = ca.jo(rwfg, type = "trace", K = 2, ecdet = "trend")
rgfg.eigen = ca.jo(rgfg, type = "eigen", K = 2, ecdet = "trend")
rgfg.trace = ca.jo(rgfg, type = "trace", K = 2, ecdet = "trend")

rfg.eigen  = ca.jo(rfg, type = "eigen", K = 2, ecdet = "trend")
rfg.trace  = ca.jo(rfg, type = "trace", K = 2, ecdet = "trend")

rhrg.eigen  = ca.jo(rhrg, type = "eigen", K = 2, ecdet = "trend")
rhrg.trace  = ca.jo(rhrg, type = "trace", K = 2, ecdet = "trend")
```

```{r}
tail(rw)
```

```{r}
rw.xts = as.xts(rw)
```

```{r}
x = ts(NA, start = c(2003, 1), end = c(2020, 3), frequency = 12)
```

```{r}
dum = ts(rep(1, 13), start=c(2008, 10), frequency = 12)
```

```{r}
rw.eigen = ca.jo(rw.xts, type = "eigen", K = 2, 
                 ecdet = "trend", season = 12)
```

```{r}
summary(rw.eigen)
```

```{r}
summary(rh.trace)
```

```{r}
cbind(rh.trace@teststat, rh.trace@cval)
```

```{r}
rh.trace@V
```

```{r}
A1 = matrix(c(1,0,0, 0,1,0), 
            nrow=3, ncol=2)
A2 = matrix(c(1,0,0, 0,0,1),
            nrow=3, ncol=2)
```

```{r}
summary(alrtest(z = rh.trace, A = A1, r = 1))
summary(alrtest(z = rh.trace, A = A2, r = 1))
```

```{r}
rw.trace
```

```{r}
series <- rnorm(1000)
y.st <- stats::filter(series, filter = c(0.6, -0.28), method = "recursive")
ar2.st <- arima(y.st, c(2,0,0), include.mean = FALSE,
                transform.pars = FALSE, method = "ML")

polyroot(c(1, -ar2.st$coef))

Mod(polyroot(c(1, -ar2.st$coef)))
Im(polyroot(c(1, -ar2.st$coef)))
Re(polyroot(c(1, -ar2.st$coef)))
```
