---
jupyter:
  jupytext:
    formats: ipynb,Rmd
    text_representation:
      extension: .Rmd
      format_name: rmarkdown
      format_version: '1.2'
      jupytext_version: 1.4.1
  kernelspec:
    display_name: R
    language: R
    name: ir
---

# โหลด packages

```{r}
source("../R/oaeTS.R", encoding = "UTF-8")
year_now = 2563
```

```{r}
price = readRDS('../data/rice/rice_price_list.rds')

# price = readRDS('../data/rice/rice_price_filled.rds')
# colnames(price) = str_replace(colnames(price), "_0", "")
# price_xts = price %>% as.xts()
# colnames(price)
```

```{r}
# head(price)
```

```{r}
# rh   = price[, c('RHFG', 'RHWS', 'RHXB')] %>% log() %>% na.omit()
# rw   = price[, c('RWFG', 'RWWS', 'RWXB')] %>% log() %>% na.omit()
# rg   = price[, c('RGFG', 'RGWS', 'RGXB')] %>% log() %>% na.omit()

# rhfg = price[, c('RHFG_NR', 'RHFG_BR', 'RHFG_SI', 'RHFG_SU', 'RHFG_UR')] %>% log() %>% na.omit()
# rwfg = price[, c('RWFG_NS', 'RWFG_KP', 'RWFG_PC', 'RWFG_PS', 'RWFG_SH')] %>% log() %>% na.omit()
# rgfg = price[, c('RGFG_UN', 'RGFG_SN', 'RGFG_KK', 'RGFG_MS', 'RGFG_KL')] %>% log() %>% na.omit()

# rfg  = price[, c('RHFG', 'RWFG', 'RGFG')] %>% log() %>% na.omit()
```

```{r}
# window(price$rh[, 1:3], start = 2004)
```

```{r}
x = list()

# vertical
x$rh = list(endo = price$rh[, 1:3] %>% window(start = 2004), 
            exo  = price$rh[, 4:5] %>% window(start = 2004))
x$rw = list(endo = price$rw[, 1:3] %>% window(start = 2004), 
            exo  = price$rw[, 4:5] %>% window(start = 2004))
x$rg = list(endo = price$rg[, 1:3] %>% window(start = 2004), 
            exo  = price$rg[, 4:5] %>% window(start = 2004))

# spatial
x$rhfg = list(endo = price$rhfg[, 1:5], exo = price$rhfg[, 6:7])
x$rwfg = list(endo = price$rwfg[, 1:5], exo = price$rwfg[, 6:7])
x$rgfg = list(endo = price$rgfg[, 1:5], exo = price$rgfg[, 6:7])

# cross products
x$rfg = list(endo = price$rfg[, 1:3], exo = price$rfg[, 4:5])
```

**Steps**
1. Model Specification
  - AR Order or Lag Selection
  - Cointegrating Rank
  - Choice of Deterministic Term
  - Testing Restrictions to alphas and betas
  - Testing Restriction for SR params and fitting subset models
2. Model Checking
  - Descriptive Analysis
  - Diagnostic Tests
    * Serial Autocorrelation
    * Heteroskedasticity
    * Normality
  - Stability Analysis
3. Analysis
  - Interpret
  - Test economic theory
  - Forecast
  - Granger Causality
  - IRF
  - FEVD


# Vertical PT: Hommali Rice

```{r}
endo = x$rh$endo
exo = x$rh$exo
```

```{r}
summary(endo %>% exp())
```

```{r}
set_figsize(10, 6)
plot.ts(endo)
```

```{r}
ur.df(endo[, 1], 'trend', lags = 12, selectlags = 'AIC') %>% summary()
```

```{r}
ur.df(endo[, 1], 'trend', lags = 12, selectlags = 'AIC')
ur.df(endo[, 2], 'trend', lags = 12, selectlags = 'AIC')
ur.df(endo[, 3], 'trend', lags = 12, selectlags = 'AIC')
ur.df(endo[, 1], 'trend', lags = 12, selectlags = 'AIC')@cval
```

```{r}
ur.df(endo[, 1], 'drift', lags = 12, selectlags = 'AIC')
ur.df(endo[, 2], 'drift', lags = 12, selectlags = 'AIC')
ur.df(endo[, 3], 'drift', lags = 12, selectlags = 'AIC')
ur.df(endo[, 1], 'drift', lags = 12, selectlags = 'AIC')@cval
```

```{r}
ur.df(endo[, 1], 'none', lags = 12, selectlags = 'AIC')
ur.df(endo[, 2], 'none', lags = 12, selectlags = 'AIC')
ur.df(endo[, 3], 'none', lags = 12, selectlags = 'AIC')
ur.df(endo[, 1], 'none', lags = 12, selectlags = 'AIC')@cval
```

## Model Selection

```{r}
VARselect(endo, lag.max = 4, type = 'none', exogen = exo)
```

```{r}
VARselect(endo, lag.max = 4, type = 'const', exogen = exo)
```

```{r}
VARselect(endo, lag.max = 4, type = 'both', exogen = exo)
```

```{r}
vecm_rh.trace = ca.jo(
  x$rh$endo, type = 'trace', ecdet = 'trend', K = 2, 
  spec = 'transitory', dumvar = x$rh$exo
)

vecm_rh.eigen = ca.jo(
  x$rh$endo, type = 'eigen', ecdet = 'trend', K = 2, 
  spec = 'transitory', dumvar = x$rh$exo
)

summary(vecm_rh.trace)
summary(vecm_rh.eigen)
```

r = 2

```{r}
vecm_rh = cajorls(vecm_rh.trace, r = 2)
```

```{r}
args(VAR)
```

```{r}
var1 = VAR(x$rh$endo , p = 1 , type = 'both', exogen = x$rh$exo)
var2 = VAR(x$rh$endo , p = 2 , type = 'both', exogen = x$rh$exo)
```

```{r}
serial.test(var2, lags.pt = 12, type = 'PT.asymptotic')
```

```{r}
arch.test(var2, lags.multi = 12)
```

```{r}
normality.test(var2)
```

```{r}
var_rh = vec2var(vecm_rh.trace)
serial.test(var_rh, lags.pt = 12, type = 'PT.asymptotic')
arch.test(var_rh, lags.multi = 12)
normality.test(var_rh)
```

## Test Law of One Price

```{r}
head(x$rh$endo)
```

```{r}
data(Canada)
var.2c <- VAR(Canada, p = 2, type = "const")
causality(var.2c, cause = "e")
```

```{r}
H1 = ca.jo(
  x$rh$endo, type = 'trace', ecdet = 'none', K = 2, 
  spec = 'transitory', dumvar = x$rh$exo
)
H1 %>% summary()
```

```{r}
H3 = matrix(c(1, -1, -1), nrow = 3, ncol = 1)

# H5 = matrix(c(1, -1, -1))
# Hlop = matrix(c(1, -1, -1), nrow = 3, ncol = 1)

bh5lrtest(z = H1, H = H3, r = 2) %>% summary()
```

```{r}
mat_lop = matrix(c(1, -1, 0, 0), 
                 nrow = 4, ncol = 1)

blrtest(z = vecm_rh.trace, H = mat_lop, r = 2)
```

```{r}
options(
  repr.plot.width=10, 
  repr.plot.height=10, 
  repr.plot.res = 300
)
```

```{r}
plot(irf(var_rh, impulse = "RHFG")) 
```

# Horizontal : Hommali Rice

```{r}
H1 = ca.jo(
  x$rwfg$endo[, c(1,2,3,4,5)], type = 'trace', ecdet = 'none', K = 2, 
  spec = 'transitory', dumvar = x$rwfg$exo
)
H1 %>% summary()
```

```{r}
HA = matrix(c( 1, 1, 1, 1,
              -1, 0, 0, 0,
               0,-1, 0, 0,
               0, 0,-1, 0,
               0, 0, 0,-1
             ), 
            ncol = 4, byrow = TRUE)

blrtest(z = H1, H = HA, r = 4) %>% summary()

# H5 = matrix(c( 1, -1), 
#             ncol = 1, byrow = TRUE)

# H5 = c(1, -1)
# Hlop = matrix(c(1, -1, -1), nrow = 3, ncol = 1)

# bh5lrtest(z = H1, H = H5, r = 1) %>% summary()
```

```{r}

```

```{r}
data(denmark)
sjd <- denmark[, c("LRM", "LRY", "IBO", "IDE")]
sjd.vecm <- ca.jo(sjd, ecdet="const", 
                  type="trace", K=2, 
                  spec="transitory", season=4)
HD0 <- matrix(c(1, -1, -1, -1, -1), c(5,1))
blrtest(sjd.vecm, H=HD0, r=1)
# summary(blrtest(sjd.vecm, H=HD0, r=1))
```

```{r}

```

```{r}
rh.cajolst = cajolst(x$rh$endo, trend = TRUE)
slotNames(rh.cajolst)
summary(rh.cajolst)
```

```{r}

```

```{r}
index(x$rh$endo)[59]
```

```{r}
rh.cajolst@bp
```

```{r}
rh.cajolst
```

```{r}
args(ca.jo)
```

```{r}
ca.jo(rh05, type = 'eigen', ecdet = "trend") %>% summary()
```

```{r}
summary(rh.cajolst)
```

```{r}
index(rw)[13]
```

# VAR


## Model Specification

```{r}
args(VARselect)
```

```{r}
VARselect(rh05, lag.max = 12, type = "both")
```

## Modle Estimation

```{r}
args(VAR)
```

```{r}
rh05.VAR = VAR(rh05, p = 2, type = "both")
```

```{r}
rh10.VAR = VAR(rh10, p = 2, type = "both")
```

```{r}
summary(rh05.VAR)
```

## Model Daignostic

```{r}
roots(rh05.VAR)
```

### Autocorrelation

```{r}
args(serial.test)
```

```{r}
rh05.VAR.ser = serial.test(rh05.VAR)
rh05.VAR.ser
```

```{r}

```

```{r}
set_figsize(10, 8)
rh05.VAR.ser = serial.test(rh05.VAR)
plot(rh05.VAR.ser, names = "RHFG")
plot(rh05.VAR.ser, names = "RHWS")
plot(rh05.VAR.ser, names = "RHXB")
```

### Hetero

```{r}
sp_vector <- c(1973.1,1961.53,1993.23,1937.27,2044.57,2054.27,2028.18,2082.2,2079.99,2094.86,2111.94,
               2099.29,2094.14,2039.87,1944.41,2024.81,2080.62,2054.08,1918.6,1904.42,2021.95,
               2075.54,2065.55,2083.89,2148.9,2170.95,2157.69,2143.02,2164.99,2246.63,2275.12,
               2329.91,2366.82,2359.31,2395.35,2433.99)
sp_ts <- ts(sp_vector,start=c(2014,7),frequency=12)
```

```{r}
dum = ts(rep(1, 13), start=c(2008, 10),frequency = 12)
```

```{r}
c(as.xts(rh05), as.xts(dum))
```

```{r}
data = replace_na(ts.union(rh05, dum), 0)
```

```{r}
# data
```

```{r}
VAR(data, p = 3, type = "both", exogen = dum)
```

```{r}

```

```{r}
rh05.VAR = VAR(rh05, p = 4, type = "both", season = 12)
```

```{r}
rh05.VAR.arch = arch.test(rh05.VAR, lags.multi = 12)
rh05.VAR.arch
```

```{r}
rh05.VAR %>% normality.test %>% plot
```

## Normality

```{r}
rh05.VAR %>% stability %>% plot
```

# Cointegration Analysis

```{r}
args(ca.jo)
```

```{r}
vars::VARselect(rh05)
vars::VARselect(rh06)
```

```{r}
rhrg  = price[, c('RHFG', 'RGFG')] %>% log() %>% na.omit()
```

```{r}
rh.eigen = ca.jo(rh, type = "eigen", K = 2, ecdet = "trend")
rh.trace = ca.jo(rh, type = "trace", K = 2, ecdet = "trend")
rw.eigen = ca.jo(rw, type = "eigen", K = 2, ecdet = "trend")
rw.trace = ca.jo(rw, type = "trace", K = 2, ecdet = "trend")
rg.eigen = ca.jo(rg, type = "eigen", K = 2, ecdet = "trend")
rg.trace = ca.jo(rg, type = "trace", K = 2, ecdet = "trend")

rhfg.eigen = ca.jo(rhfg, type = "eigen", K = 2, ecdet = "trend")
rhfg.trace = ca.jo(rhfg, type = "trace", K = 2, ecdet = "trend")
rwfg.eigen = ca.jo(rwfg, type = "eigen", K = 2, ecdet = "trend")
rwfg.trace = ca.jo(rwfg, type = "trace", K = 2, ecdet = "trend")
rgfg.eigen = ca.jo(rgfg, type = "eigen", K = 2, ecdet = "trend")
rgfg.trace = ca.jo(rgfg, type = "trace", K = 2, ecdet = "trend")

rfg.eigen  = ca.jo(rfg, type = "eigen", K = 2, ecdet = "trend")
rfg.trace  = ca.jo(rfg, type = "trace", K = 2, ecdet = "trend")

rhrg.eigen  = ca.jo(rhrg, type = "eigen", K = 2, ecdet = "trend")
rhrg.trace  = ca.jo(rhrg, type = "trace", K = 2, ecdet = "trend")
```

```{r}
tail(rw)
```

```{r}
rw.xts = as.xts(rw)
```

```{r}
x = ts(NA, start = c(2003, 1), end = c(2020, 3), frequency = 12)
```

```{r}
dum = ts(rep(1, 13), start=c(2008, 10), frequency = 12)
```

```{r}
rw.eigen = ca.jo(rw.xts, type = "eigen", K = 2, 
                 ecdet = "trend", season = 12)
```

```{r}
summary(rw.eigen)
```

```{r}
summary(rh.trace)
```

```{r}
cbind(rh.trace@teststat, rh.trace@cval)
```

```{r}
rh.trace@V
```

```{r}
A1 = matrix(c(1,0,0, 0,1,0), 
            nrow=3, ncol=2)
A2 = matrix(c(1,0,0, 0,0,1),
            nrow=3, ncol=2)
```

```{r}
summary(alrtest(z = rh.trace, A = A1, r = 1))
summary(alrtest(z = rh.trace, A = A2, r = 1))
```

```{r}
rw.trace
```

```{r}

```
